Clause Size Reduction with \alluip Learning

### 2020-05-15 paper.tex
fix the last typo, hopefully
last typo
fix font, fit into the page limit

\documentclass[runningheads]{llncs}
\usepackage{times}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{xspace}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{array}
\usepackage{xcolor}
\usepackage{theorem}
\usepackage[T1]{fontenc}
\usepackage{hyperref}

@@

\begin{figure} 
\begin{center}
\begin{tabular}{ | m{3.2cm} | m{3.8cm}| m{2cm} | m{2.75cm} |  } 
\hline
Solver & \# solved (SAT, UNSAT) & PAR-2 & avg. clause Size \\ 
\hline
\oneuip & 221 (132, 89)  & 5018.89 & 62.6  \\ 
\hline
$\allUipPure$ &\textbf{228} (135, \bf{93}) +7 & 4867.37 & 49.88 \\
\hline
$\allUipMin$ & 226 (135, 91) +5 & 4890.67 & 45.2 \\ 
\hline
$\allUipAct$ & 226 (135, 91)  +5 & \textbf{4866.94} & 47.7 \\
\hline
$\allUipIn$ & 225 (\textbf{138}, 87) +4 & 4958.49 & 52.12 \\
\hline
$\allUipEx$ & 223 (134, 89) +2& 5015.23& \textbf{43.2} \\
\hline
\end{tabular}
\end{center}
\caption{Benchmark results of $\MapleBase$ with \oneuip, $\allUipPure$, $\allUipMin$, $\allUipAct$,
$\allUipIn$, and $\allUipEx$ on SAT2019 race main track.}
\label{fig:t4}
\end{figure}

@@

The results in Fig.~\ref{fig:t4}, ~\ref{fig:len_pdf} and
\ref{fig:len_compare} indicate $\allUipMin$ often achieves higher
clause reduction than $\allUipPure$. We also observed that
$\allUipMin$ attempted algorithm~\ref{alg:i-uip} more frequently than
$\allUipPure$ (28.8\% vs 16.1\%), and is more likely to succeed
(59.3\% vs 43.4\%). This observation agrees with our experiment
results.

@@

  \begin{minipage}[t]{0.5\textwidth}
   \includegraphics[width=\textwidth]{figures/clause_reduction_PDF.pdf}
   \caption{Relative clause size reduction distribution. The $X$ axis
      indicates the relative size of difference between all-UIP and
      1-UIP clauses (calculated as
      $\dfrac{|C_1|-|C_i|}{|C_1|}$ ) for
      each instance, and the $Y$ axis shows the probability density.}
       \label{fig:len_pdf}
  \end{minipage}
  & &
  \begin{minipage}[t]{0.5\textwidth}
    \includegraphics[width=\textwidth]{figures/clause_size_compare.pdf}
    \caption{Average clause size comparison plot. Each point in the
      plot represents an instance. The $X$ and $Y$ axes shows the clause
      length from $\stablealluip$ and \oneuip, respectively. Each green (red)
      dot represents an compared instance between $\MapleBase$ and
      $\MapleIUIMin$ ($\allUipPure$).}
      \label{fig:len_compare}
  \end{minipage}

\end{tabular}
}
\end{figure}

\subsubsection{Reduced Proof Sizes with $\stablealluip$.}
A learning scheme that yields smaller clauses (lemmas) might also
construct smaller causal proofs. For 88 UNSAT instances solved mutually
by $\allUipPure$, $\allUipMin$ and \oneuip schemes, we additionally
compare the size of the optimized DRAT proof from the three learning schemes. 
We used the DRAT-trim tool
\cite{DBLP:conf/sat/WetzlerHH14} with a 30000 second timeout to check and
optimize every DRAT proof once \footnote{Applying DRAT-trim multiple
  times can further reduce the proof size until a fix-point. However,
  the full optimization is time consuming and was impractical for our
  experiments.}.

The average optimized DRAT proof from $\allUipMin$ and $\allUipPure$
are 556.6MB and 698.5MB, respectively. Both sizes are significantly
smaller than the average optimized proof size from \oneuip, 824.9MB.
The average proof size reduction per instance for $\allUipMin$ and
$\allUipPure$ was 16.5\% and 3.6\% compared with \oneuip, which roughly
correlate with our observed clause size observation in
Fig~\ref{fig:len_pdf}.

\begin{figure} 
\begin{center}
\fss{8pt}{10pt}
\begin{tabular}{|l|l|l|l|} 
\hline
Solver & \#solved (SAT, UNSAT) $\Delta$ & PAR-2 & avg. clause Size \\ 
\hline
SAT 2017 Winner & & & \\
$\MapleSeven$ & 232 (135, 97) & 4755.96 & 61.9  \\ 
\hline
$\MapleSeven$-all-pure & \textbf{244 (146, 98)} +12 & \textbf{4504.18} & 43.76 \\
\hline
$\MapleSeven$-all-min & 240 (144, 96) +8 & 4601.25 & \textbf{36.97} \\ 
\hline
$\MapleSeven$-all-act & 237 (140, 97) +5 & 4678.434 & 43.62 \\ 
\hline
$\MapleSeven$-all-inclusive & 234 (137, 97) +2 & 4718.03 & 37.96 \\
\hline
\hline
SAT 2018 Winner & & & \\
$\MapleEightShort$ & 236 (138, 98) & 4671.81 & 61.69 \\
\hline
$\MapleEightShort$-all-pure & \textbf{241} (\textbf{142, 99}) +5 & \textbf{4598.18} & 44.19 \\
\hline
$\MapleEightShort$-all-min & 236 (141, 95) +0 & 4683.92 & 38.05 \\ 
\hline
$\MapleEightShort$-all-act & 240 (141, \textbf{99}) +4 & 4626.99 & 41.16 \\
\hline
$\MapleEightShort$-all-inclusive & 240 (\textbf{142}, 98) +4 & 4602.13 & \textbf{37.52} \\
\hline
\hline
SAT 2019 Winner & & & \\
$\MapleNineShort$ & 238 (140, \textbf{98}) & 4531.24 & 60.91 \\
\hline
$\MapleNineShort$-all-pure & 240 (142, \textbf{98}) +2 & 4519.08 &  43.32\\
\hline
$\MapleNineShort$-all-min & 244 (\textbf{148}, 96) +6 & \textbf{4419.84} & \textbf{36.88} \\
\hline
$\MapleNineShort$-all-act & 243 (146, 97) + 5 & 4476.73 & 40.65 \\
\hline
$\MapleNineShort$-all-inclusive & \textbf{243} (\textbf{148}, 95) +5 & 4455.76 & 37.02 \\
\hline
\hline
SAT 2019 Competitor & & &\\
$\expSATShort$ & 237 (137, 100)  & 4628.96 & 63.19 \\
\hline
$\expSATShort$-all-pure & 235 (136, 99) $-$2  & 4668.96 & 48.26 \\
\hline
$\expSATShort$-all-min & 241 (143, 98) +4 & 4524.28 & 46.29 \\ 
\hline
$\expSATShort$-all-act & 244 (143, \textbf{101}) +7 & \textbf{4460.92} & 47.25 \\
\hline
$\expSATShort$-all-inclusive & \textbf{245} (\textbf{146}, 99) +8 & 4475.76 & \textbf{45.33}
\\
\hline
\hline
 & & &\\
$\defaultcadical$ & 249 (150, 99)  & \textbf{4311.76} & 101.96 \\
\hline
$\defaultcadical$-all-pure & 248 (151, 97) $-$1  & 4373.38 & 82.44  \\
\hline
$\defaultcadical$-all-min & 248 (149, 99) $-$1 & 4398.31 & 43.93 \\ 
\hline
$\defaultcadical$-all-act & \textbf{252} (\textbf{152}, \textbf{100}) +3 & 4331.56 & 47.34 \\
\hline
$\defaultcadical$-all-inclusive & 251 (\textbf{153}, 98) +2 & 4335.88 & \textbf{42.61}
\\
\hline
\end{tabular}
\end{center}
\caption{Benchmark results of \oneuip, $\allUipPure$. $\allUipMin$, $\allUipAct$ and $\allUipIn$ on SAT2019 race main track.}
\label{fig:t5}
\end{figure}

@@

On the SAT 2017 race winner $\MapleSeven$, all four configurations of
$\stablealluip$ outperformed \oneuip learning by 12, 8, 5, and 2 instances,
respectively. $\allUipPure$ solved more UNSAT and SAT instances while
the other configurations improved on solving SAT instances. The clause
size reduction of $\stablealluip$ is more significant on $\MapleSeven$
than on to $\MapleBase$.

@@

On the SAT 2019 race winner $\MapleNineShort$, all four
$\stablealluip$ configurations solved instances 2, 6, 5 and 5 more instances
than \oneuip, respectively. $\MapleNineShort$ prioritizes clauses that
are learned multiple times. We observed that clauses learnt from
$\stablealluip$ configurations are less likely to be duplicated. As an
example, $\allUipMin$ in average, added 12\% less duplicated clauses
into the core clause database than \oneuip. This observation is
surprising, and the cause is unclear.

On $\expSATShort$, three out of four $\stablealluip$ configurations solved
4, 7 and 8 more instances than \oneuip learning. We noticed that
$\allUipAct$ and $\allUipIn$ show significant better performance than
$\allUipMin$ and $\allUipPure$ on this solver. The random walk
simulation branching heuristic, however, didn't impact the performance
of $\stablealluip$ schemes significantly.

$\defaultcadical$ with \oneuip solved 249 total instances. 
Applying $\allUipAct$ and $\allUipIn$ helped the solver solve 2 and 3 more instance, respectively.  
The average learnt clause size of \oneuip on 
$\defaultcadical$ is much larger than other solvers' (101 vs 60) 
while the average clause size of $\stablealluip$ configurations are similar on all solvers.

>>>>

\documentclass[runningheads]{llncs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{xspace}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{array}
\usepackage{xcolor}
\usepackage{theorem}
\usepackage[T1]{fontenc}
\usepackage{hyperref}

@@

\begin{figure} 
\begin{center}
\begin{tabular}{ | m{3.2cm} | m{3.8cm}| m{2cm} | m{2.75cm} |  } 
\hline
Solver & \# solved (SAT, UNSAT) & PAR-2 & avg. clause Size \\ 
\hline
\oneuip & 221 (132, 89)  & 5018.89 & 62.6  \\ 
\hline
$\allUipPure$ &\textbf{228} (135, \bf{93}) +7 & 4867.37 & 49.88 \\
\hline
$\allUipMin$ & 226 (135, 91) +5 & 4890.67 & 45.2 \\ 
\hline
$\allUipAct$ & 226 (135, 91)  +5 & \textbf{4866.94} & 47.7 \\
\hline
$\allUipIn$ & 225 (\textbf{138}, 87) +4 & 4958.49 & 52.12 \\
\hline
$\allUipEx$ & 223 (134, 89) +2& 5015.23& \textbf{43.2} \\
\hline
\end{tabular}
\end{center}
\caption{Results of $\MapleBase$ with \oneuip, $\allUipPure$, $\allUipMin$, $\allUipAct$,
$\allUipIn$, and $\allUipEx$ on SAT2019 race main track.}
\label{fig:t4}
\end{figure}

@@

The results in Fig.~\ref{fig:t4}, \ref{fig:len_pdf} and
\ref{fig:len_compare} indicate $\allUipMin$ often achieves higher
clause reduction than $\allUipPure$. We also observed that
$\allUipMin$ attempted algorithm~\ref{alg:i-uip} more frequently than
$\allUipPure$ (28.8\% vs 16.1\%), and is more likely to succeed
(59.3\% vs 43.4\%). This observation agrees with our experiment
results.

@@

  \begin{minipage}[t]{0.5\textwidth}
   \includegraphics[width=\textwidth]{figures/clause_reduction_PDF.pdf}
   \caption{Relative clause size reduction distribution. The $X$ axis
      indicates the relative size of difference between all-UIP and
      1-UIP clauses (calculated as
      $\dfrac{|C_1|-|C_i|}{|C_1|}$ ) for
      each instance, and the $Y$ axis shows the probability density.}
       \label{fig:len_pdf}
  \end{minipage}
  & &
  \begin{minipage}[t]{0.5\textwidth}
    \includegraphics[width=\textwidth]{figures/clause_size_compare.pdf}
    \caption{Average clause size comparison plot. Each point in the
      plot represents an instance. The $X$ and $Y$ axes shows the clause
      length from $\stablealluip$ and \oneuip, respectively. Each green (red)
      dot represents an compared instance between $\MapleBase$ and
      $\MapleIUIMin$ ($\allUipPure$).}
      \label{fig:len_compare}
  \end{minipage}
\end{tabular}
}
\end{figure}

\subsubsection{Reduced Proof Sizes with $\stablealluip$.}
A learning scheme that yields smaller clauses (lemmas) might also
construct smaller causal proofs. For 88 UNSAT instances solved mutually
by $\allUipPure$, $\allUipMin$ and \oneuip schemes, we additionally
compared the size of the optimized DRAT proof from the three learning schemes. 
We used the DRAT-trim tool
\cite{DBLP:conf/sat/WetzlerHH14} with a 30000 second timeout to check and
optimize every DRAT proof once\footnote{Applying DRAT-trim multiple
  times can further reduce the proof size until a fix-point. However,
  the full optimization is too time consuming for our
  experiments.}.

The average optimized DRAT proof from $\allUipMin$ and $\allUipPure$
are 556.6MB and 698.5MB, respectively. Both sizes are significantly
smaller than the average optimized proof size from \oneuip, 824.9MB.
The average proof size reduction per instance for $\allUipMin$ and
$\allUipPure$ is 16.5\% and 3.6\% against \oneuip, which roughly
correlate with our clause size observation in
Fig~\ref{fig:len_pdf}.

\begin{figure} 
\begin{center}
\fss{8pt}{10pt}
\begin{tabular}{|l|l|l|l|} 
\hline
Solver & \#solved (SAT, UNSAT) $\Delta$ & PAR-2 & avg. clause Size \\ 
\hline
SAT 2017 Winner & & & \\
$\MapleSeven$ & 232 (135, 97) & 4755.96 & 61.9  \\ 
\hline
$\MapleSeven$-all-pure & \textbf{244 (146, 98)} +12 & \textbf{4504.18} & 43.76 \\
\hline
$\MapleSeven$-all-min & 240 (144, 96) +8 & 4601.25 & \textbf{36.97} \\ 
\hline
$\MapleSeven$-all-act & 237 (140, 97) +5 & 4678.434 & 43.62 \\ 
\hline
$\MapleSeven$-all-inclusive & 234 (137, 97) +2 & 4718.03 & 37.96 \\
\hline
\hline
SAT 2018 Winner & & & \\
$\MapleEightShort$ & 236 (138, 98) & 4671.81 & 61.69 \\
\hline
$\MapleEightShort$-all-pure & \textbf{241} (\textbf{142, 99}) +5 & \textbf{4598.18} & 44.19 \\
\hline
$\MapleEightShort$-all-min & 236 (141, 95) +0 & 4683.92 & 38.05 \\ 
\hline
$\MapleEightShort$-all-act & 240 (141, \textbf{99}) +4 & 4626.99 & 41.16 \\
\hline
$\MapleEightShort$-all-inclusive & 240 (\textbf{142}, 98) +4 & 4602.13 & \textbf{37.52} \\
\hline
\hline
SAT 2019 Winner & & & \\
$\MapleNineShort$ & 238 (140, \textbf{98}) & 4531.24 & 60.91 \\
\hline
$\MapleNineShort$-all-pure & 240 (142, \textbf{98}) +2 & 4519.08 &  43.32\\
\hline
$\MapleNineShort$-all-min & 244 (\textbf{148}, 96) +6 & \textbf{4419.84} & \textbf{36.88} \\
\hline
$\MapleNineShort$-all-act & 243 (146, 97) + 5 & 4476.73 & 40.65 \\
\hline
$\MapleNineShort$-all-inclusive & \textbf{243} (\textbf{148}, 95) +5 & 4455.76 & 37.02 \\
\hline
\hline
SAT 2019 Competitor & & &\\
$\expSATShort$ & 237 (137, 100)  & 4628.96 & 63.19 \\
\hline
$\expSATShort$-all-pure & 235 (136, 99) $-$2  & 4668.96 & 48.26 \\
\hline
$\expSATShort$-all-min & 241 (143, 98) +4 & 4524.28 & 46.29 \\ 
\hline
$\expSATShort$-all-act & 244 (143, \textbf{101}) +7 & \textbf{4460.92} & 47.25 \\
\hline
$\expSATShort$-all-inclusive & \textbf{245} (\textbf{146}, 99) +8 & 4475.76 & \textbf{45.33}
\\
\hline
\hline
$\defaultcadical$ & 249 (150, 99)  & \textbf{4311.76} & 101.96 \\
\hline
$\defaultcadical$-all-pure & 248 (151, 97) $-$1  & 4373.38 & 82.44  \\
\hline
$\defaultcadical$-all-min & 248 (149, 99) $-$1 & 4398.31 & 43.93 \\ 
\hline
$\defaultcadical$-all-act & \textbf{252} (\textbf{152}, \textbf{100}) +3 & 4331.56 & 47.34 \\
\hline
$\defaultcadical$-all-inclusive & 251 (\textbf{153}, 98) +2 & 4335.88 & \textbf{42.61}
\\
\hline
\end{tabular}
\end{center}
\caption{Benchmark results of \oneuip, $\allUipPure$. $\allUipMin$, $\allUipAct$ and $\allUipIn$ on SAT2019 race main track.}
\label{fig:t5}
\end{figure}

@@

On the SAT 2017 race winner $\MapleSeven$, all four configurations of
$\stablealluip$ outperformed \oneuip learning by 12, 8, 5, and 2 instances,
respectively. $\allUipPure$ solved more UNSAT and SAT instances while
the other configurations improved on solving SAT instances. The clause
size reduction of $\stablealluip$ is more significant on this solver
than on to $\MapleBase$.

@@

On the SAT 2019 race winner $\MapleNineShort$, all four
$\stablealluip$ configurations solved instances 2, 6, 5 and 5 more instances
than \oneuip, respectively. $\MapleNineShort$ prioritizes clauses that
are learned multiple times. We observed that
$\stablealluip$ clauses are less likely to be duplicated. As an
example, $\allUipMin$ in average, added 12\% less duplicated clauses
into the core clause database than \oneuip. This observation is
surprising, and the cause is unclear.

On $\expSATShort$, three out of four $\stablealluip$ configurations solved
4, 7 and 8 more instances than \oneuip learning. We noticed that
both $\allUipAct$ and $\allUipIn$ show better performance than
$\allUipMin$ and $\allUipPure$ on this solver. The random walk
simulation branching heuristic, however, didn't impact the performance
of $\stablealluip$ schemes significantly.

$\defaultcadical$ with \oneuip solved 249 total instances. 
Applying $\allUipAct$ and $\allUipIn$ helped the solver solve 3 and 2 more instances, respectively.  
The average learnt clause size of \oneuip on 
$\defaultcadical$ is much larger than other solvers' (101 vs 60) 
while the average clause size of $\stablealluip$ configurations are similar on all solvers.
